---
title: "Independent data analysis"
author: "Harsimran Ghai"
date: "2026-03-02"
output: html_document
---
 Github link: https://github.com/harsimranghai91-eng/CHEP-898-Independent-Analysis-Part-1-.git
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Identify a dataset to work on 

I am interested in myocardial infarction and therefore I choose a dataset containing risk factors for myocardial infarction (heart attack) from kaggle website. This dataset is based on chinese population. 

The link to the datset is as follows: https://www.kaggle.com/datasets/ankushpanday2/heart-attack-risk-dataset-of-china

# Gaols of the analysis

My goal using this dataset is as follows:

Develop the best model to predict the heart attack using the below mentioned predictor. I will be using random forest and binary logistic regression, as my outcome variable is binary in nature.


# Statistical Analysis Plan

The stages of analysis are as follows

## Data Wrangling
  -Loading the dataset and examine the structure of variables
  -Checking for missing and duplicate data
  -Correct the formatting of the data set

## Descriptive Analysis
  - Calculating the measure of central tendancy and measure of dispersion for continuous data.
  - Calculating frequency and proportion for categorical variables
  
## Data Visualisation
  - Bar Chart
  - Box Plots

## Binary Logistic Regression
  - Splitting the data into testing (70%) and training data (30%)
  - Building a model using the dependent variable that is heart attack (Binary in nature:0/1) and the predictors variables (see below).
  - Checking the model performance on the testing data (Cross validation).

## Classifiication Tree
Classification tree model will be built using the below mentioned predictors to predictor heart attack.

Both models will be compared by creating the confusion matrix and calculating the associated measures (Sensitivity, specificity, accuracy, etc.) to select the best model.
# Variable Information

## Dependent Variable

1. **Heart attack**  
    Binary: "Yes" = no heart attack, "No" = heart attack

## Independent Variables

1. **Age** – Continuous
2. **Blood Pressure** – Continuous
3. **CVD Risk Score** – Continuous
4. **Hypertension** – Binary
5. **Diabetes** – Binary
6. **Obesity** – Binary
7. **Cholesterol** – Categorical (High, Normal, Low)
8. **Air Pollution Exposure** – Categorical (High, Medium, Low)
9. **Physical Activity** – Categorical (High, Medium, Low)
10. **Diet Score** – Categorical (Poor, Moderate, Healthy)
11. **Stress Level** – Categorical (Low, Medium, High)
12. **Alcohol Consumption** – Binary
13. **Family History CVD** – Binary
14. **Healthcare Access** – Categorical (Poor, Moderate, Good)
15. **Rural or Urban** – Binary
16. **Hospital Availability** – Categorical (Low, Medium, High)
17. **TCM Use** – Binary
18. **Employment Status** – Categorical (Employed, Unemployed, Retired)
19. **Education Level** – Categorical (None, Primary, Secondary, Higher)
20. **Income Level** – Categorical (Low, Middle, High)
21. **Chronic Kidney Disease** – Binary
22. **Previous Heart Attack** – Binary

# Data Wrangling

```{r}
# Loading the data
data<-read.csv("heart_attack_china.csv")
```

```{r}
# Loading the Packages
library(tidyverse)
library(tidymodels)
library(reportRmd)
library(sjPlot)
library(plotly)
library(psych)
library(parallel)
library(finalfit)
library(gtsummary)
library(mlbench)
library(vip)
library(rsample)
library(tune)
library(recipes)
library(yardstick)
library(parsnip)
library(glmnet)
library(themis)
library(corrr)
library(performance)
library(utils)
library(see)
library(forestmodel)
```

# Taking a look into data set

```{r}
glimpse(data)
```

This dataset has a total of 28 variables and a sample size of 239,266

# Checking duplicates

```{r}
sum(duplicated(data))
```

No duplicated observation were observed in this dataset.

# Checking the missing data

```{r}
any(is.na(data))
```
No missing data was found in this data set, so we don't need to do anything for the missing data.

# Checking the structure in the data set

```{r}
str(data)
```

# Changing the Structure of variables

```{r}
data <- data%>%mutate(across(c( Gender,
                                Smoking_Status,
                                Hypertension,
                                Diabetes,
                                Obesity,
                                Cholesterol_Level,
                                Air_Pollution_Exposure,
                                Physical_Activity,
                                Diet_Score,
                                Stress_Level,
                                Alcohol_Consumption,
                                Family_History_CVD,
                                Healthcare_Access,
                                Rural_or_Urban,
                                Hospital_Availability,
                                TCM_Use,
                                Employment_Status,
                                Education_Level,
                                Income_Level,
                                Chronic_Kidney_Disease,
                                Previous_Heart_Attack,
                                Heart_Attack
                            
),~ as.factor(.)))
```

# Descriptive Analysis

```{r}
rm_covsum(
  data = data,
                        covs = c('Age',
                                'Blood_Pressure',
                                'CVD_Risk_Score',
                                'Hypertension',
                                'Diabetes',
                                'Obesity',
                                'Cholesterol_Level',
                                'Air_Pollution_Exposure',
                                'Physical_Activity',
                                'Diet_Score',
                                'Stress_Level',
                                'Alcohol_Consumption',
                                'Family_History_CVD',
                                'Healthcare_Access',
                                'Rural_or_Urban',
                                'Hospital_Availability',
                                'TCM_Use',
                                'Employment_Status',
                                'Education_Level',
                                'Income_Level',
                                'Chronic_Kidney_Disease',
                                'Previous_Heart_Attack',
                                'Heart_Attack')
)
```

# Bar Graphs
```{r}
ggplot(data, aes(x = Diabetes, fill = Heart_Attack)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Diabetes by Disease",
    x = "Diabetes",
    y = "Count",
    fill = "Heart_Attack"
  ) +
  theme_minimal()
```

```{r}
ggplot(data, aes(x =Hypertension , fill = Heart_Attack)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Hypertension by Disease",
    x = "Hypertension",
    y = "Count",
    fill = "Heart_Attack"
  ) +
  theme_minimal()
```

```{r}
ggplot(data, aes(x =Obesity , fill = Heart_Attack)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Obesity by Disease",
    x = "Obesity",
    y = "Count",
    fill = "Heart_Attack"
  ) +
  theme_minimal()
```

```{r}
ggplot(data, aes(x =Cholesterol_Level , fill = Heart_Attack)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Cholesterol_Level by Disease",
    x = "Cholesterol_Level",
    y = "Count",
    fill = "Heart_Attack"
  ) +
  theme_minimal()
```

```{r}
ggplot(data, aes(x = Air_Pollution_Exposure, fill = Heart_Attack)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Air_Pollution_Exposure by Disease",
    x = "Air_Pollution_Exposure",
    y = "Count",
    fill = "Heart_Attack"
  ) +
  theme_minimal()
```

```{r}
ggplot(data, aes(x = Physical_Activity, fill = Heart_Attack)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Physical_Activity by Disease",
    x = "Physical_Activity",
    y = "Count",
    fill = "Heart_Attack"
  ) +
  theme_minimal()
```

```{r}
ggplot(data, aes(x = Diet_Score, fill = Heart_Attack)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Diet_Score by Disease",
    x = "Diet_Score",
    y = "Count",
    fill = "Heart_Attack"
  ) +
  theme_minimal()
```

```{r}
ggplot(data, aes(x = Stress_Level, fill = Heart_Attack)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Stress_Level by Disease",
    x = "Stress_Level",
    y = "Count",
    fill = "Heart_Attack"
  ) +
  theme_minimal()
```

```{r}
ggplot(data, aes(x = Alcohol_Consumption, fill = Heart_Attack)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Alcohol_Consumption by Disease",
    x = "Alcohol_Consumption",
    y = "Count",
    fill = "Heart_Attack"
  ) +
  theme_minimal()
```

```{r}
ggplot(data, aes(x = Family_History_CVD, fill = Heart_Attack)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Family_History_CVD by Heart_Attack",
    x = "Family_History_CVD",
    y = "Count",
    fill = "Heart_Attack"
  ) +
  theme_minimal() 
```

```{r}
ggplot(data, aes(x = Healthcare_Access, fill = Heart_Attack)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Healthcare_Access by Heart_Attack",
    x = "Healthcare_Access",
    y = "Count",
    fill = "Heart_Attack"
  ) +
  theme_minimal()
```

```{r}
ggplot(data, aes(x = Rural_or_Urban, fill = Heart_Attack)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Rural_or_Urban by Heart_Attack",
    x = "Rural_or_Urban",
    y = "Count",
    fill = "Heart_Attack"
  ) +
  theme_minimal() 
```

```{r}
ggplot(data, aes(x = Hospital_Availability, fill = Heart_Attack)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Hospital_Availability by Heart_Attack",
    x = "Hospital_Availability",
    y = "Count",
    fill = "Heart_Attack"
  ) +
  theme_minimal() 
```

```{r}
ggplot(data, aes(x = TCM_Use, fill = Heart_Attack)) +
  geom_bar(position = "dodge") +
  labs(
    title = "TCM_Use by Heart_Attack",
    x = "TCM_Use",
    y = "Count",
    fill = "Heart_Attack"
  ) +
  theme_minimal() 
```

```{r}
ggplot(data, aes(x = Employment_Status, fill = Heart_Attack)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Employment_Status by Heart_Attack",
    x = "Employment_Status",
    y = "Count",
    fill = "Heart_Attack"
  ) +
  theme_minimal() 
```

```{r}
ggplot(data, aes(x = Education_Level, fill = Heart_Attack)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Education_Level by Heart_Attack",
    x = "Education_Level",
    y = "Count",
    fill = "Heart_Attack"
  ) +
  theme_minimal() 
```

```{r}
ggplot(data, aes(x = Income_Level, fill = Heart_Attack)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Income_Level by Heart_Attack",
    x = "Income_Level",
    y = "Count",
    fill = "Heart_Attack"
  ) +
  theme_minimal() 
```

```{r}
ggplot(data, aes(x = Chronic_Kidney_Disease, fill = Heart_Attack)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Chronic_Kidney_Disease by Heart_Attack",
    x = "Chronic_Kidney_Disease",
    y = "Count",
    fill = "Heart_Attack"
  ) +
  theme_minimal() 
```

```{r}
ggplot(data, aes(x = Heart_Attack, y = Age)) +
  geom_boxplot() +
  labs(
    x = "Heart Attack Status",
    y = "Age",
    title = "Distribution of Age by Heart Attack Status"
  ) +
  theme_minimal()
```

```{r}
ggplot(data, aes(x = Heart_Attack, y = Blood_Pressure)) +
  geom_boxplot() +
  labs(
    x = "Heart Attack Status",
    y = "Blood Pressure",
    title = "Distribution of Blood Pressure by Heart Attack Status"
  ) +
  theme_minimal()
```

```{r}
ggplot(data, aes(x = Heart_Attack, y = CVD_Risk_Score)) +
  geom_boxplot() +
  labs(
    x = "Heart Attack Status",
    y = "CVD Risk Scores",
    title = "Distribution of CVD Risk Scores by Heart Attack Status"
  ) +
  theme_minimal()
```
